Exploring Local Explanations of Non-linear Models Using Animated Linear Projections

### 2022-01-16 _paper.rmd
- add 2 sequences to the SHAP_bd plot - sequences of featuers rather than permutations.

<!-- Paper structure -->
The remainder of this paper is organized as follows. The following Section, \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced. Section \@ref(sec:tour) explains the animations of continuous linear projections. Section \@ref(sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure. Then Section \@ref(#sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features. We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explored in the future.

@@

<!-- tree SHAP -->
For our application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018]. Instead of aggregating over permutations of the feature order, tree SHAP approximates instance-level feature importance by exploring the structure of the decision trees. Tree SHAP is only compatible with tree-based models; we illustrate random forests. The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.

```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills. A star offensive and defensive player are compared, L. Messi and V. van Dijk, respectively. Panel (a) is the break-down plot showing one additive feature-attribution. Panel (b) shows the distribution of attribution for each feature across 25 permutations of predictors, with the mean displayed as a dot, for each players. Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
knitr::include_graphics("./figures/shap_distr_bd.png")
```

>>>>

<!-- Paper structure -->
The remainder of this paper is organized as follows. The following Section, \@ref(sec:explanations) covers the background of the local explanation and the traditional visuals produced. Section \@ref(sec:tour) explains the animations of continuous linear projections. Section \@ref(sec:applicationdesign) discusses the visual layout in the interactive interface, how they facilitate analysis, data preprocessing, and package infrastructure. Then Section \@ref(sec:casestudies) illustrates the application to supervised learning with categorical and quantitative response features. We conclude with Section \@ref(sec:cheemdiscussion) of the insights gained and directions that might be explored in the future.

@@

```{r shapdistrbd, echo=F, fig.align='center', out.width="60%", fig.cap = "Illustration of the use of SHAP values for a random forest model for salaries of FIFA 2020 players based on nine predictors corresponding to different skills. A star offensive and defensive player are compared, L. Messi and V. van Dijk, respectively. Panel (a) shows breakdown plot of three seqences of the features, order and magnitude change. Panel (b) shows the distribution of attribution for each feature across 25 seqences of predictors, with the mean displayed as a dot, for each players. Offense and movement are important for Messi but not van Dijk, and conversely, defense and power are important for van Dijk but not Messi."}
knitr::include_graphics("./figures/shap_distr_bd.png")
```

<!-- tree SHAP -->
For our application, we use _tree SHAP_, a variant of SHAP that enjoys a lower computational complexity [@lundberg_consistent_2018]. Instead of aggregating over sequences of the features, tree SHAP calculates instance-level feature importance by exploring the structure of the decision trees. Tree SHAP is only compatible with tree-based models; we illustrate random forests. The following section will use normalized SHAP values as a projection basis (call this the _attribution projection_) which will have coefficients varied in order to further scrutinize the feature contributions.
